---
title: "COVID Data Analysis"
author: "Zhen"
date: "2023-12-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, message=FALSE, warning=FALSE}
# install.packages("tidyverse")
library(tidyverse)
library(lubridate)
```

## Import Data

```{r}
us_cases <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")

global_cases <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")

us_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")

global_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

```

## Check Data

```{r}

us_cases
global_cases
us_deaths
global_deaths

```

## Transform Data

Transform data to be useable.

```{r}
library(lubridate)
global_cases <- global_cases %>%
    pivot_longer(cols = -c('Province/State', 'Country/Region', Lat, Long), 
                 names_to = "date",
                 values_to = "cases") %>%
    select(-c(Lat,Long))
    
    
global_cases  
```

```{r}
global_deaths <- global_deaths %>%
    pivot_longer(cols = -c('Province/State', 'Country/Region', Lat, Long), 
                 names_to = "date",
                 values_to = "deaths") %>%
    select(-c(Lat,Long))

global_deaths
```

```{r}
us_cases <- us_cases %>%
    pivot_longer(cols = -(UID:Combined_Key),
                 names_to = "date",
                 values_to = "cases") %>%
    select(Admin2:cases) %>%
    mutate(date = mdy(date)) %>%
    select(-c(Lat, Long_))

us_cases
```

```{r}
us_deaths <- us_deaths %>%
    pivot_longer(cols = -(UID:Population),
                 names_to = "date",
                 values_to = "deaths") %>%
    select(Admin2:deaths) %>%
    mutate(date = mdy(date)) %>%
    select(-c(Lat, Long_))

us_deaths
```

## Summary& Clean Data

```{r}
summary(us_cases)
summary(us_deaths)
summary(global_cases)
summary(global_deaths)
```

```{r}
library(lubridate)
global <- global_cases %>%
  full_join(global_deaths) %>%
  rename(Country_Region =`Country/Region`,
         Province_State = 'Province/State') %>%
  mutate(date = mdy(date))

global
```

```{r}

us <- us_cases %>%
  full_join(us_deaths)

us
```

```{r}
summary(us)
summary(global)
```

From the summary, we found the cases and deaths column in us table have negative number, which should be positive number.

```{r}
us <- us %>%
  group_by(Province_State,Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(death = abs(deaths), case = abs(cases)) %>%
  select(Province_State, Country_Region, date, case, death, Population) %>%
  ungroup()

us
```

Add a column for death/case rate.

```{r}
us <- us %>%
  group_by(Province_State,Country_Region, date) %>%
  summarize(cases = sum(case), deaths = sum(death), Population = sum(Population)) %>%
  mutate(death_case_rate = deaths/cases) %>%
  select(Province_State, Country_Region, date, cases, deaths, death_case_rate,Population) %>%
  ungroup()

us
```

```{r}
global <- global %>%
  group_by(Province_State,Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths)) %>%
  mutate(death_case_rate = deaths/cases) %>%
  select(Province_State, Country_Region, date, cases, deaths, death_case_rate) %>%
  ungroup()

global
```

```{r}
summary(us)
summary(global)

```

## Visualiazations

Create liner graphs for the cases over time.

```{r}
usa <- us %>%
  group_by(Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(death_case_rate = deaths/cases) %>%
  select(Country_Region, date, cases, deaths, death_case_rate,Population) %>%
  ungroup()
 
```

```{r}
globals <- global %>%
  group_by(date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths)) %>%
  mutate(death_case_rate = deaths/cases) %>%
  select(date, cases, deaths, death_case_rate) %>%
  ungroup()

globals
```

```{r}
usa %>%
  ggplot(aes(x= date, y = cases)) +
  geom_point(aes(color = "cases")) + 
  scale_y_log10() + 
   theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "US COVID Case", y = NULL)+
  guides(fill = FALSE)
```

```{r}
globals %>%
  ggplot(aes(x= date, y = cases)) +
  geom_point(aes(color = "cases")) + 
  scale_y_log10() + 
   theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "Global COVID Case", y = NULL)+
  guides(fill = FALSE)
```

From the graph, we found that the number of COVID-19 cases in global increased the most between March and June 2020, then gradually declined and leveled off around February 2022; the number of COVID-19 cases in global increased the most between February and May 2020, then gradually declined and leveled off around February 2022.

Create liner graphs for the deaths over time.

```{r}
usa %>%
  ggplot(aes(x= date, y = deaths)) +
  geom_point(aes(color = "deaths")) + 
  scale_y_log10() + 
   theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "US COVID Deaths", y = NULL)+
  guides(fill = FALSE)
```

```{r}
globals %>%
  ggplot(aes(x= date, y = deaths)) +
  geom_point(aes(color = "deaths")) + 
  scale_y_log10() + 
   theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "Global COVID Deaths", y = NULL)+
  guides(fill = FALSE)
```

From the graph, we found that the number of COVID-19 deaths in global increased the most between February and Apirl 2020, then gradually declined and leveled off around June 2021; the number of COVID-19 deaths in global increased the most between February and March 2020, then gradually declined and leveled off around May 2020.

Create a liner graph for death/case rate in the US.

```{r}
usa %>%
  ggplot(aes(x= date, y = death_case_rate)) +
  geom_point(aes(color = "death case rate")) + 
  scale_y_log10() + 
   theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90)) +
  labs(title = "US COVID Death/Case Rate", y = NULL)+
  guides(fill = FALSE)
```

From the graph we can tell that the death/case rate was very high at the beginning because we didn't have a cure for COVID-19. And it continued to decline after April 2020.I think it's because of the invention of the COVID-19 treatment drug.

## Summary

From this data, we can tell:

Global COVID-19 Incidence Trends:
The global number of COVID-19 cases experienced its most substantial surge between March and June 2020, followed by a gradual decline and stabilization by February 2022. Additionally, there was another notable increase in cases between February and May 2020, with a subsequent leveling off by February 2022.

Global COVID-19 Mortality Patterns:
The peak in global COVID-19-related deaths occurred between February and April 2020, showing a gradual decline and stabilization around June 2021. Similarly, another spike in deaths was observed between February and March 2020, followed by a leveling off around May 2020.

Death/Case Ratio and Treatment Advancements:
The initial high death-to-case ratio, attributed to the absence of a cure for COVID-19, began to decline after April 2020. This decline is likely linked to advancements in the development and availability of COVID-19 treatment drugs, indicating a positive impact on mortality rates.
